TrueSkill Through Time: the \proglang{Julia}, \proglang{Python} and \proglang{R} packages

### 2021-03-29 article.tex
TTT: autores
TTT: nueva figura optimización, y discusión de la figura ATP

%% - \author{} with primary affiliation
%% - \Plainauthor{} without affiliations
%% - Separate authors by \And or \AND (in \author) or by comma (in \Plainauthor).
%% - \AND starts a new line, \And does not.
\author{Gustavo Landfried \\Universidad de Buenos Aires
   \And Matias Mazzanti \\Universidad de Buenos Aires
   \And Esteban Mocskos \\Universidad de Buenos Aires}
\Plainauthor{Gustavo Landfried, Matias Mazzanti, Esteban Mocskos }

@@

\en{In this last example we use the entire history of matches of the Association of Tennis Professionals (ATP).}
\es{En este último ejemplo utilizamos toda la historia de partidas de la Asociación de Tenistas Profesionales (ATP).}
%
\en{The database has \num{447000} games starting at year 1915 until 2020 with more than \num{19000} participating players and is publicly available\footnote{A zip file containing a csv file per year can be downloaded from \nolinkurl{https://github.com/JeffSackmann/tennis_atp}}.}
\es{La base de datos cuenta con \num{447000} mil partidas que van desde el año 1915 hasta el 2020, en las que participan \num{19000} jugadores.}
%
\en{The information is formatted and stored in a single \texttt{csv} file\footnote{Available at \nolinkurl{https://github.com/glandfried/tennis_atp/releases/download/atp/history.csv.zip}} including both single and double matches: if the column \texttt{double} has the letter \texttt{t}, the game is a double match.}
\es{La información contenida en los archivos \texttt{csv} contiene tanto partidas simples como dobles, indicado por la columna \texttt{double}.}
%
\en{Each game has a unique identifier (i.e. \texttt{match\_id}) and tournament's round number (i.e. \texttt{round\_number}) where 0 represents the final game, 1 the semi-final, and so on.}
\es{Tenemos un identificador por la partida (\texttt{match\_id}) y el número de ronda (\texttt{round\_number}) donde 0 representa la partida final, 1 la semi-final, etc.}
%
\en{The file also contains four columns for the players' identifiers in the game (the last two players are null in the case of a single): for example \texttt{w2\_id} is the identifier of the second player of the winning team and \texttt{l1\_name} is the name of the first player of the losing team.}
\es{Contiene cuatro columnas correspondiente a los identificadores de los jugadores y cuatro columnas correspondiente a sus nombres, por ejemplo \texttt{w2\_id} es el identificador del segundo jugador del equipo ganador (los segundos jugadores tienen valores nulos en partidas simples) y \texttt{l1\_name} es el nombre del primer jugador del equipo perdedor.}
%
\en{Finally, we have the name of the tournament (\texttt{tour\_name}), its identifier (\texttt{tour\_id}), the start date of the tournament (\texttt{time\_start}) and the type of surface (\texttt{ground}).}
\es{Finalmente tenemos el nombre del torneo (\texttt{tour\_name}), su identificado (\texttt{tour\_id}), la fecha de inicio del torneo (\texttt{time\_start}) y el tipo de piso (\texttt{ground}).}
%
\en{Here we only show the \proglang{Julia} version because it is more efficient than \proglang{Python} and \proglang{R} versions, which can be also found in the appendix \ref{sec:appendix_atp_code}.}
\es{Aquí mostramos solo la versión de \proglang{Julia} debido a que es la más eficiente, las versiones de \proglang{Python} y \proglang{R} se pueden entonrar en el anexo \ref{sec:appendix_atp_code}.}
%
\begin{lstlisting}[backgroundcolor=\color{julia},label=lst:atp, caption={\en{The history of the Association of Tennis Professionals}\es{Historia sintética}}, belowskip=0.1cm, aboveskip=0.1cm]
using CSV; using Dates
data = CSV.read("atp.csv")

@@

h = History(composition=composition, times = times, sigma = 2.0, gamma = 0.02)
convergence(h,epsilon=0.01, iterations=10)
\end{lstlisting}
%
\en{In this code we open the file \texttt{atp.csv}, create the variables \texttt{times} and \texttt{composition}, and instantiate the class \texttt{History}.}
\es{En este código abrimos el archivo \texttt{csv}, creamos las variables \texttt{times} y \texttt{composition} e instanciamos la clase \texttt{History}.}
%
\en{Temporal batches are defined as days elapsed since a reference date to the tournament start.}
\es{Los lotes temporales se definen como días transcurridos desde una fecha de referencia hasta el inicio del torneo.}
%
\en{When generating the list \texttt{composition} we discriminate whether the games are double or single based on the column \texttt{double}.}
\es{Al generar la lista \texttt{composition} discriminamos si los juegos son dobles o individuales en base a la columna \texttt{double}.}
%
\en{The results are implied by the order in which we set the composition, always placing the winning team first.}
\es{Los resultados quedan impl\'icitos por el orden en que armamos la composición, poniendo siempre primero al equipo ganador.}
%
\en{When initializing the class \texttt{History} we also modify the default values of \texttt{sigma} and \texttt{gamma} based on an optimization procedure previously performed (more details next).}
\es{Al inicializar la clase \texttt{History} modificamos los valores por defecto de \texttt{sigma} y \texttt{gamma}, basados en un procedimiento de optimización realizado previamente (más detalles a continuación).}
%
\en{Finally, we use the \texttt{convergence()} method to obtain TrueSkill Through Time estimates explicitly selecting the convergence criterion: when the change between iterations is less than $0.01$ or when $10$ iterations are performed.}
\es{Finalmente, convergemos la historia para obtener las estimaciones de TrueSkill Through Time, indicando explícitamente el criterio de corte: cuando el cambio entre iteraciones sea menor a $0.01$, o cuando se alcanzan las $10$ iteraciones.}
%
\en{Figure~\ref{fig:atp} presents the learning curves of eight famous players of the ATP's history.}
\es{En la figura~\ref{fig:atp} graficamos las curvas de aprendizaje de ocho jugadores famosos de la historia de la ATP.}
%
\begin{figure}[ht!]
    \centering
    \includegraphics[page={1},width=.7\linewidth]{figures/atp}
    \caption{ATP}
    \label{fig:atp}
\end{figure}

\todo[inline]{discusion}
% 
% \en{The parameter \texttt{epsilon} indicates the convergence condition, the maximum change in all posteriors before and after an iteration.}
% \es{El parámetro \texttt{epsilon} indica la condici\'on de convergencia, el cambio máximo en todos los posteriors antes y despu\'es de una iteraci\'on.}
% %
% \en{Finally, the number of iterations \texttt{iter} functions as a cut-off condition in case convergence has not been reached.}
% \es{Por \'ultimo, la cantidad de iteraciones \texttt{iter} funciona como condici\'on de corte en caso de no haber alcanzado la convergencia.}

% \en{Immediately after the history is initialized, a summary of the data is printed.}
% \es{Inmediatamente despu\'es de que se inicializa la historia, se imprime un resumen de los datos.}
%
% \en{In this case, 3 events, 3 batches and 3 agents are reported.}
% \es{En este caso se reportan 3 eventos, 3 batches y 3 agentes.}
%
% \en{Batches are used to group events that happened at the same time.}
% \es{Los batches sirven para agrupar eventos que ocurrieron al mismo tiempo.}
%
% \en{We will see an example with batches later on.}
% \es{Ya veremos un ejemplo con batches más adelante.}
%
% \en{As we do not indicate the time of each game, the system places the games in different batches.}
% \es{Como no le indicamos el tiempo de cada partida, el sistema ubica las partida en batches distintos.}

>>>>

%% - \author{} with primary affiliation
%% - \Plainauthor{} without affiliations
%% - Separate authors by \And or \AND (in \author) or by comma (in \Plainauthor).
%% - \AND starts a new line, \And does not.
\author{Gustavo Landfried \\Universidad de Buenos Aires
   \And Esteban Mocskos \\Universidad de Buenos Aires}
\Plainauthor{Gustavo Landfried, Esteban Mocskos }

@@

\en{In this last example we use the entire history of matches of the Association of Tennis Professionals (ATP).}
\es{En este último ejemplo utilizamos toda la historia de partidas de la Asociación de Tenistas Profesionales (ATP).}
%
\en{The database has \num{447000} games starting at year 1915 until 2020 with more than \num{19000} participating players and is publicly available\footnote{Source data at \nolinkurl{https://github.com/JeffSackmann/tennis_atp}}.}
\es{La base de datos cuenta con \num{447000} mil partidas que van desde el año 1915 hasta el 2020, en las que participan \num{19000} jugadores.}
%
\en{The information is formatted and stored in a single \texttt{csv} file\footnote{Available at \nolinkurl{https://github.com/glandfried/tennis_atp/releases/download/atp/history.csv.zip}} including both single and double matches: if the column \texttt{double} has the letter \texttt{t}, the game is a double match.}
\es{La información contenida en los archivos \texttt{csv} contiene tanto partidas simples como dobles, indicado por la columna \texttt{double}.}
%
\en{Each game has a unique identifier (i.e. \texttt{match\_id}) and tournament's round number (i.e. \texttt{round\_number}) where 0 represents the final game, 1 the semi-final, and so on.}
\es{Tenemos un identificador por la partida (\texttt{match\_id}) y el número de ronda (\texttt{round\_number}) donde 0 representa la partida final, 1 la semi-final, etc.}
%
\en{The file also contains four columns for the players' identifiers in the game (the last two players are null in the case of a single): for example \texttt{w2\_id} is the identifier of the second player of the winning team and \texttt{l1\_name} is the name of the first player of the losing team.}
\es{Contiene cuatro columnas correspondiente a los identificadores de los jugadores y cuatro columnas correspondiente a sus nombres, por ejemplo \texttt{w2\_id} es el identificador del segundo jugador del equipo ganador (los segundos jugadores tienen valores nulos en partidas simples) y \texttt{l1\_name} es el nombre del primer jugador del equipo perdedor.}
%
\en{Finally, we have the name of the tournament (\texttt{tour\_name}), its identifier (\texttt{tour\_id}), the start date of the tournament (\texttt{time\_start}) and the type of surface (\texttt{ground}).}
\es{Finalmente tenemos el nombre del torneo (\texttt{tour\_name}), su identificado (\texttt{tour\_id}), la fecha de inicio del torneo (\texttt{time\_start}) y el tipo de piso (\texttt{ground}).}
%
\en{Here we only show the \proglang{Julia} version because it is more efficient than \proglang{Python} and \proglang{R} versions, which can be also found in the appendix \ref{sec:appendix_atp_code}.}
\es{Aquí mostramos solo la versión de \proglang{Julia} debido a que es la más eficiente, las versiones de \proglang{Python} y \proglang{R} se pueden entonrar en el anexo \ref{sec:appendix_atp_code}.}
%
\begin{lstlisting}[backgroundcolor=\color{julia},label=lst:atp, caption={\en{The history of the Association of Tennis Professionals}\es{Historia sintética}}, belowskip=0.1cm, aboveskip=0.1cm]
using CSV; using Dates
data = CSV.read("atp.csv")

@@

h = History(composition=composition, times = times, sigma = 1.6, gamma = 0.036)
convergence(h,epsilon=0.01, iterations=10)
\end{lstlisting}
%
\en{In this code we open the file \texttt{atp.csv}, create the variables \texttt{times} and \texttt{composition}, and instantiate the class \texttt{History}.}
\es{En este código abrimos el archivo \texttt{csv}, creamos las variables \texttt{times} y \texttt{composition} e instanciamos la clase \texttt{History}.}
%
\en{Temporal batches are defined as days elapsed since a reference date to the tournament start.}
\es{Los lotes temporales se definen como días transcurridos desde una fecha de referencia hasta el inicio del torneo.}
%
\en{When generating the list \texttt{composition} we discriminate whether the games are double or single based on the column \texttt{double}.}
\es{Al generar la lista \texttt{composition} discriminamos si los juegos son dobles o individuales en base a la columna \texttt{double}.}
%
\en{The results are implied by the order in which we set the composition, always placing the winning team first.}
\es{Los resultados quedan impl\'icitos por el orden en que armamos la composición, poniendo siempre primero al equipo ganador.}
%
\en{When initializing the class \texttt{History} we set the values of \texttt{sigma} and \texttt{gamma} based on an optimization procedure previously performed (Figure~\ref{fig:optimization}).}
\es{Al inicializar la clase \texttt{History} modificamos los valores de \texttt{sigma} y \texttt{gamma} basados en un procedimiento de optimización realizado previamente (figura~\ref{fig:optimization}).}
%
\en{Finally, we use \texttt{convergence()} method to obtain TrueSkill Through Time estimates explicitly selecting the convergence criterion: when the change between iterations is less than $0.01$ or when $10$ iterations are performed.}
\es{Finalmente, usamos el método \texttt{convergence()} para obtener las estimaciones de TrueSkill Through Time, indicando explícitamente el criterio de corte: cuando el cambio entre iteraciones sea menor a $0.01$, o cuando se alcanzan las $10$ iteraciones.}
%
\en{Figure~\ref{fig:atp} presents the learning curves of eight famous players of the ATP's history.}
\es{En la figura~\ref{fig:atp} graficamos las curvas de aprendizaje de ocho jugadores famosos de la historia de la ATP.}
%
\begin{figure}[ht!]
    \centering
    \includegraphics[page={1},width=.7\linewidth]{figures/atp}
    \caption{
    \en{Estimated learning curves of 8 famous players in ATP history.}
    \es{Estimación de las curvas de aprendizaje de 8 jugadores famosos de la historia de la ATP.}
    %
    \en{The shaded area represents an uncertainty equivalent to one standard deviation.}
    \es{El área sombrada representa una incertidumbre equivalente a un desvío estandar.}
    }

    \label{fig:atp}
\end{figure}
%
\en{The advantage of TrueSkill Through Time, over TrueSkill, is that by modeling the entire dynamic process it allows us to make comparisons of skills over time.}
\es{La ventaja de TrueSkill Through Time, respecto de TrueSkill, es que al modelar el proceso dinámico permite hacer comparaciones temporales de la habilidad.}
%
\en{It is interesting to see that the skill of tennis players did not increase so much over the years: on the contrary the players of the 1980s were more skilled than those of the 1990s, and reached a skill similar to what Federer, Nadal and Djokovic had in 2020.}
\es{Es interesante ver que la habilidad de los tenistas no aumentó tanto con los años: al contrario los jugadores de la década de los 80 fueron más habilidosos que aquellos de la década de los 90, y alcanzaron una habilidad similar a la que Federer, Nadal y Djokovic tuvieron en el año 2020.}
%
\en{Perhaps the most notable change introduced by professionalism is the stability of learning curves, which is seen in today's players as opposed to the early players.}
\es{Quizás el cambio más notable introducido por el profesionalismo sea la estabilidad de las curvas de aprendizaje, que se aprecia en los jugadores actuales a diferencia de los primeros jugadores.}
